---
title: "A1_HYDE_20261425, Mari Hyde"
output:
  html_document:
    df_print: paged
date: "2025-01-07"
---

Repository: https://github.com/marihyde/Biol432_A1


The first section of code opens the updated measurements file, which includes calculated limb volumes. 
head function is used to display a sample of the table. 
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)

Measurements<-read.csv("Measurements")
head(Measurements)
```



Dplyr is used to order the rows in Measurements by Organism, observer, then limb volume, using the arrange function. 
```{r}
Sort_Species<- Measurements %>% arrange(Organism, Observer, Limb_volume)
head(Sort_Species)
```


This section groups the data by each species using the group_by function. Then each volume is calculated using the mean function, and printed to display the average limb volume for each species. 
```{r}
Average_Volume<- Sort_Species %>% group_by(Organism) %>% summarize(Avg_Volume= mean(Limb_volume, na.rm=TRUE))
print(Average_Volume)
```
This chunk sorts the data by observer and species using the group_by function. the Summarise function is counting the number of organisms viewed by each observer for each species, and displaying them as "ObservationCount" in the table. 
```{r}
number_observations<- Sort_Species %>% group_by(Observer, Organism) %>% summarise(ObservationCount=n(), .groups="drop")
print(number_observations)
```


This chunk displays a boxplot showing the distribution of limb volumes across different species. 
```{r}
ggplot(Measurements, aes(x= Organism, y=Limb_volume))+
  geom_boxplot()+
  theme_light()+
  labs(title= "Limb Volume as a Function of Species", 
       x="Species",
       y="Limb Volume")
```

This chunk of code converted the limb volume to elongated measurements, making it easier to see the distribution in the next section. 
```{r}
elongated_measurements<- Measurements %>% gather(key="Measurement_Type", value="limbvolume", Limb_volume)
head(elongated_measurements)
```

This chunk is displaying the distribution of the elongated limb volume measurements for each species on a histogram. The function facet_wrap splits the data between species, for multiple histograms. 
```{r}
ggplot(elongated_measurements, aes(x=limbvolume))+
  geom_histogram(bins=20, fill= "lightpink", color="black", alpha=1)+
  facet_wrap(~Organism)+
  theme_light()+
  labs(title="Distribution of Limb Volume across Dog Species", x="Limb Volume", y="")
```


